---
title: "Task1"
author: "Artem"
date: "14/10/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 title

Question description Atomic Types

Create the following variables related to your chosen object:

4 numeric values (e.g., population, temperature, size, rating)
3 integer values (e.g., number of moons, buildings, inhabitants)
2 logical values (e.g., has_water, is_inhabited)
3 character values (e.g., name, region, category)

My object is Kiev city

```{r}
# code for solving task
#numeric


temperature <- 1.15
population <- 2.904346
size <- 8.398470
cars <- 9.242323

#integer


skyscrapers <- 1222
inhabitants <- 2483789
shopingsmalls <- 78

#logical


river <- TRUE
ocean <- FALSE


#character


name <- "OleksandrUsyk"
region <- "Kiev"
name <- "Shevchenko"






## Question 1.2.Vectors


# 1. Numeric vector (mix of numeric and integer values)
numeric_vector <- c(temperature, population, size, cars)
numeric_vector

# 2. Sequence vector from 25 to 100 with a step of 5
sequence_vector <- seq(from = 25, to = 100, by = 5)
sequence_vector

# 3. Logical vector of length 10 with random TRUE/FALSE values
set.seed(7)  # set seed for reproducibility
logical_vector <- sample(c(TRUE, FALSE), size = 10, replace = TRUE)
logical_vector


## Question 1.3.Matrix


# Create a 3x3 matrix with random integers between -50 and 50
set.seed(10)  # for reproducibility
env_matrix <- matrix(sample(-50:50, 9 ), nrow = 3, ncol = 3)

env_matrix

rowMeans(env_matrix)
colMeans(env_matrix)






```
## Question 1.4. List

Question description

```{r}
# code for solving task
my_list <- list(population, skyscrapers, river, name, numeric_vector, sequence_vector, logical_vector, env_matrix)
my_list
```
## Question 5 Factor

Question description

Create a factor variable representing categories within your universe, e.g.:

```{r}
# code for solving task

my_factor <- factor(c("big", "big", "small", "medium", "small"), levels = c("small", "medium", "big"))
my_factor
```
## Question 6 Data Frame

Question description

Create a data frame related to your object that includes:

5–8 observations (rows)
At least 3 columns of different data types
(e.g., name, numeric value, logical flag, category)
Then:

Rename the columns using names() or colnames().
Display its structure with str() and its summary with summary().
Add a new column derived from the others (for example, a ratio or condition).

```{r}
data <- data.frame(name=c("centr","chreshatik", "ziloti", "piyanavishnya", "maka_scena"), numeric = c(16.4, 14.6, 25.7, 25.1, 18.2), logical = c(T, F, F, T, T), category = c("big", "small", "medium", "medium", "small"))

colnames(data) <- c("name", "size", "has_mall", "size")

data

str(data)

summary(data)

data$size_ratio <- data$size / max(data$size)

data

```


