---
title: "Task1"
author: "Artem"
date: "14/10/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 title

Question description Atomic Types

Create the following variables related to your chosen object:

4 numeric values (e.g., population, temperature, size, rating)
3 integer values (e.g., number of moons, buildings, inhabitants)
2 logical values (e.g., has_water, is_inhabited)
3 character values (e.g., name, region, category)

My object is Kiev city

```{r}
# code for solving task
#numeric


temperature <- 1.15
population <- 2.904346
size <- 8.398470
cars <- 9.242323

#integer


skyscrapers <- 1222
inhabitants <- 2483789
shopingsmalls <- 78

#logical


river <- TRUE
ocean <- FALSE


#character


name <- "OleksandrUsyk"
region <- "Kiev"
name <- "Shevchenko"






## Question 1.2.Vectors


# 1. Numeric vector (mix of numeric and integer values)
numeric_vector <- c(temperature, population, size, cars)
numeric_vector

# 2. Sequence vector from 25 to 100 with a step of 5
sequence_vector <- seq(from = 25, to = 100, by = 5)
sequence_vector

# 3. Logical vector of length 10 with random TRUE/FALSE values
set.seed(7)  # set seed for reproducibility
logical_vector <- sample(c(TRUE, FALSE), size = 10, replace = TRUE)
logical_vector


## Question 1.3.Matrix


# Create a 3x3 matrix with random integers between -50 and 50
set.seed(10)  # for reproducibility
env_matrix <- matrix(sample(-50:50, 9 ), nrow = 3, ncol = 3)

env_matrix

rowMeans(env_matrix)
colMeans(env_matrix)






```
## Question 1.4. List

Question description

```{r}
# code for solving task
my_list <- list(population, skyscrapers, river, name, numeric_vector, sequence_vector, logical_vector, env_matrix)
my_list
```
## Question 5 Factor

Question description

Create a factor variable representing categories within your universe, e.g.:

```{r}
# code for solving task

my_factor <- factor(c("big", "big", "small", "medium", "small"), levels = c("small", "medium", "big"))
my_factor
```
## Question 6 Data Frame

Question description

Create a data frame related to your object that includes:

5–8 observations (rows)
At least 3 columns of different data types
(e.g., name, numeric value, logical flag, category)
Then:

Rename the columns using names() or colnames().
Display its structure with str() and its summary with summary().
Add a new column derived from the others (for example, a ratio or condition).

```{r}
data <- data.frame(name=c("centr","chreshatik", "ziloti", "piyanavishnya", "maka_scena"), numeric = c(16.4, 14.6, 25.7, 25.1, 18.2), logical = c(T, F, F, T, T), category = c("big", "small", "medium", "medium", "small"))

colnames(data) <- c("name", "size", "has_mall", "size")

data

str(data)

summary(data)

data$size_ratio <- data$size / max(data$size)

data

```
---
title: "Task2"
author: "Artem"
date: "04/11/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Task 2: Subsetting & Missing Data

Question description

Work with objects created in Task 1.

```{r}
# Recreate necessary data from Task 1

temperature <- 1.15
population <- 2.904346
size <- 8.398470
cars <- 9.242323

skyscrapers <- 1222
inhabitants <- 2483789
shopingsmalls <- 78

river <- TRUE
ocean <- FALSE

name <- "Shevchenko"
region <- "Kiev"

numeric_vector <- c(temperature, population, size, cars)
sequence_vector <- seq(from = 25, to = 100, by = 5)
set.seed(7)
logical_vector <- sample(c(TRUE, FALSE), size = 10, replace = TRUE)

set.seed(10)
env_matrix <- matrix(sample(-50:50, 9), nrow = 3, ncol = 3)

my_list <- list(population, skyscrapers, river, name, numeric_vector, sequence_vector, logical_vector, env_matrix, description = region)

data <- data.frame(
  name = c("centr","chreshatik", "ziloti", "piyanavishnya", "maka_scena"),
  size = c(16.4, 14.6, 25.7, 25.1, 18.2),
  has_mall = c(T, F, F, T, T),
  category = c("big", "small", "medium", "medium", "small")
)
```

## 2.1 Querying Vectors

```{r}
# 2.1.1 Sequence Vector
sequence_vector[3]            # Select and print 3rd element
sequence_vector[2:5]          # Select and print elements 2–5
sequence_vector[sequence_vector > 50]   # Values greater than 50

# 2.1.2 Logical Vector
logical_vector[logical_vector]   # Only TRUE values
which(logical_vector)            # Indices of TRUE values
```

## 2.2 Querying Your “Universe” List

```{r}
# Access and print the description string using $
my_list$description

# Access and print your 3x3 matrix using [[ ]]
my_list[[8]]

# Access and print the second element from numeric_vector inside the list
my_list[[5]][2]
```

## 2.3 Investigating Your Data Frame

```{r}
# Print the first 3 rows
head(data, 3)

# Print the value from 4th row, 1st column
data[4, 1]

# Print only the logical column
data$has_mall

# Logical subsetting: rows where size > 20
df_subset <- data[data$size > 20, ]
df_subset

# %in% subsetting: select small and medium
df_subset_2 <- data[data$category %in% c("small", "medium"), ]
df_subset_2
```

## 2.4 Handling “Corrupted” Data (NA)

```{r}
# Copy data frame
df_corrupted <- data

# Introduce 3 NAs
df_corrupted[2, 2] <- NA
df_corrupted[5, 3] <- NA
df_corrupted[3, 4] <- NA

df_corrupted

# Find NAs
sum(is.na(df_corrupted))
colSums(is.na(df_corrupted))

# Mean of size column (with and without NA)
mean(df_corrupted$size)
mean(df_corrupted$size, na.rm = TRUE)

# Filter out rows with NAs
df_clean <- na.omit(df_corrupted)
df_clean
```

## 2.5 Bonus Challenge: which()

```{r}
# Find the row with highest and lowest size
highest_size_index <- which.max(data$size)
data[highest_size_index, ]

lowest_size_index <- which.min(data$size)
data[lowest_size_index, ]
```


`---
title: "Task 3: Real World Dataset"
author: "Artem"
format:
  html:
    embed-resources: true
editor: source
---

## 3.1 Choose & Load Your Dataset

Dataset source: Local file uploaded by student  
File: **kiev_districts.csv**

```{r}
library(dplyr)
library(readr)

data <- read_csv("C:/Users/makuh/OneDrive/Рабочий стол/ua.csv")

glimpse(data)



# First inspection
glimpse(data)
head(data)
summary(data)
```
```{r}
data %>% 
  select(city, population) %>% 
  arrange(desc(population)) %>% 
  head(5)

```
```{r}
data %>%
group_by(admin_name) %>% 
  summarise(avg_population = mean(population, na.rm = TRUE)) %>% 
  arrange(desc(avg_population))
```
```{r}
data %>% 
  group_by(capital) %>% 
  summarise(count = n())
```
```{r}
data %>% 
  filter(!is.na(population), !is.na(population_proper)) %>% 
  filter(population_proper < population) %>% 
  select(city, population, population_proper)
```
```{r}
data %>% 
filter(lat > 50) %>% 
  select(city, lat, admin_name, population)
```



